---
title: "p8105_hw6_ob2305"
author: "Olya Besedina"
data: "11/21/2019"
output: github_document
---
  
```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(patchwork)

knitr::opts_chunk$set(
 	echo = TRUE,
 	warning = FALSE,
 	fig.width = 8, 
   fig.height = 6,
   out.width = "90%"
 )

options(
   ggplot2.continuous.colour = "viridis",
   ggplot2.continuous.fill = "viridis"
 )

 scale_colour_discrete = scale_colour_viridis_d
 scale_fill_discrete = scale_fill_viridis_d

 theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

```{r, tidy dataset}
birthweight = read.csv(file = "./data/birthweight.csv") %>% 
   janitor::clean_names() %>% 
   mutate(
      babysex = as.factor(babysex),
      frace = as.factor(frace),
      malform = as.factor(malform),
      mrace = as.factor(mrace))

# check for missing data
skimr::skim(birthweight)

birthweight_1 = birthweight %>% 
   mutate(
   wtgain = as.factor(wtgain))
```

## Fit regression model

Regression model was first fit for the outcome of interest - baby birthweight in grams - with all the variables in the `birthweight` dataset. Non-significant predictors were removed one-by-ne and model was refitted; this process was repeated untill only significant predictors were left in the model. Those predictors are: `babysex, bhead, blength, delwt, fincome, gaweeks, mheight, momage, mrace, parity, ppbmi, smoken`. 

```{r, my regression model}
my_model  = lm(bwt ~ babysex + bhead + blength + delwt + fincome + gaweeks + mheight + momage + mrace + parity + ppbmi + smoken, data = birthweight)

adj_r_sq_mm = my_model %>% 
   broom::glance() %>% 
   pull(adj.r.squared)

aic_mm = 
   my_model %>% 
   broom::glance() %>% 
   pull(AIC)
```

Model diagnostics were performed using `broom::glance`. Adjusted r-squared for the obtained model is `r adj_r_sq_mm` and AIC is `r aic_mm`


Varience is depicted by residuals and fitted values, which are distributed around 0 and show somewhat constant varience. Residuals were cut off at 1000 to exclude extreme outliers.

```{r, residuals and fitted values}
# add residuals
birthweight = modelr::add_residuals(birthweight, my_model) 
# add predictions
birthweight = modelr::add_predictions(birthweight, my_model)

birthweight%>%
  ggplot(aes(x = pred, y = resid))+
  geom_point(color = "purple", alpha = .5)+
   # cut off at 1000 to exclude extreme outliers
   ylim(-1000, 1000)
```


## Compare to two other model

```{r}
model_1 = lm(bwt ~ blength + gaweeks, data = birthweight)

model_1 %>% 
  broom::glance()

model_2 = lm(bwt ~ babysex*blength*bhead, data = birthweight)
summary(model_2)
model_2 %>% 
  broom::glance()
```

## Cross validation

Cross validation of birth weight dataset was performed 100 times using `modelr::crossv_mc`. Linear regression was fit for each sample and rmse were compared visually in violin plot. 
```{r}
cv_df = 
  crossv_mc(birthweight, 100)

# convert every list to dataframe
cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_results = 
   cv_df %>% 
   mutate(
      my_model = map(train, ~my_model),
      model_1 = map(train, ~ model_1),
      model_2 = map(train, ~ model_2),
      
      rmse_my_model = map2_dbl(my_model, test, ~rmse(model = .x, data = .y)),
      rmse_model_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
      rmse_model_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y)))
```

A violin plot depicts distribution of rmse for each model from 100 training - testing splits. Based on the plot `my_model` is the best model among the 3 models depicted, because it has the lowest rmse values. 

```{r}
cv_results %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, fill = model)) + 
  geom_violin(alpha = .5)
```

# Problem 2

```{r, load weather dataset}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
Weather dataset was obtained from NOAA via rnoaa pachage. It contains 365 observations and 6 variables `r names(weather_df)`.


Bootstraping `weather_df` 5000 times using `modelr` package.

```{r, bootstraping weather_df}
#dataset for log(b0*b1)
bootstrap_beta = 
   weather_df %>% 
   modelr::bootstrap(n = 5000) %>%
   mutate(
      models = map(strap, ~lm(tmax ~ tmin, data = .)),
      # obtain beta0 and beta1 using broom::tidy
      results = map(models, broom::tidy)
   ) %>% 
   unnest(results) %>% 
   select(.id:estimate, -models) %>% 
   pivot_wider(
      names_from = "term",
      values_from = "estimate") %>% 
   rename(intercept = '(Intercept)') %>% 
   mutate(
      log_b0_b1 = log(intercept*tmin)
   ) %>% 
   select(.id, log_b0_b1)

#dataset for r^2
bootstrap_r2 = 
   weather_df %>% 
   modelr::bootstrap(n = 5000) %>%
   mutate(
      models = map(strap, ~lm(tmax ~ tmin, data = .)),
      # obtain r^2 values using broom::glance
      results = map(models, broom::glance)
   ) %>% 
   unnest(results) %>% 
   select(.id, r.squared)

# join two datasets
bootstrap_data = 
   inner_join(bootstrap_beta, bootstrap_r2, by = ".id") %>% 
   janitor::clean_names()
```

The plot below depicts distribution for the estimates obtained from bootstraping. 

* Distribution of R-squared does ~N distribution, but is slightly right skewed. The average R-squared value for the 5000 fitted linear regression is about 47%, meaning that about 47% values of `tmax` is explained by the value of `tmin`. 
* Distribution of log(beta0*beta1) ~N distribution with the average value of 16. 

```{r}
# density plot of r^2
rsq = 
   bootstrap_data%>%
   ggplot(aes(x = r_squared))+
   geom_density(fill = "yellow", alpha = .7)+
   labs(
      title = "Distribution of R^2",
      x = "R^2",
      y = "Density",
      caption = "Data from NOAA via rnoaa pachage"
   )

# density plot of log(b0*b1)
log_b0_b1 = 
   bootstrap_data%>%
   ggplot(aes(x = log_b0_b1))+
   geom_density(fill = "pink", alpha = .7)+
   labs(
      title = "Distribution of log(beta0 * beta1)",
      x = "log(beta0 * beta1)",
      y = "Density",
      caption = "Data from NOAA via rnoaa pachage")

rsq + log_b0_b1
```

The 95% Confidence Interval for R-squared estimates is between `r quantile(bootstrap_data$r_squared, probs = 0.025)` and `r quantile(bootstrap_data$r_squared, probs = 0.975)`. 

The 95% Confidence Interval for log(beta0 * beta1) is between `r quantile(bootstrap_data$log_b0_b1, probs = 0.025)` and `r quantile(bootstrap_data$log_b0_b1, probs = 0.975)`








